Low-Cost Data Deduplication for Virtual Machine Backup in Cloud Storage

### 2013-03-08 hotstorage13/arch.tex
evaluation changes

The data flow of our partition-based  dupliate detectionis depicted in
Figure~\ref{fig:flow}. let $v$ be the total number of virtual machines hosted in  
a cluster,  we divide the backup into $k$  iterations. 
\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{steps.pdf}
\caption{ Processing steps of dupliate detection.}
\label{fig:flow}
\end{figure}

\begin{itemize}
\item
The first step of each iteration is to read  $v/k$  virtual machine images in parallel on the cloud 
cluster and issue backup requests. For those segments that have been modified since last backup indicated by 
dirty bits,  the system divides  each segment into a sequence of chunk blocks,  computes the meta 
information such as chunk fingerprints,  sends the metadata of a chunk block, and accumulate  into 
a partition bucket on the temporary disk storage. 
The partition mapping is using a hash function applied to the content fingerprint. 
Assuming all machines have a  homogenous resource configuration, each machine is evently  assigend with
$q$ partitons of content hash and accumulates corresponding requests on the disk. 
There are two options to assign buffers in each machine to buffer requests for each partition.
1) Each machine has  $p\times q$ send buffers corresponding to $p\times q$ partitions in the cluster
since any content block in a VM image of this machine can be sent to any of these partitions.
2) Each machine has $p$ receive buffers correspond to $p$ machines in the cluster and $q$ receive 
buffers to temporarly accmulate messages from other machines, and then output each buffer to the local disk
when it becomes full.  Option 2 is much more efficient than Option 1 because $p+q$ is much smaller than
$p\times q$, as a result each buffer has a bigger size to accmulate data and results in 
much less disk seek overhead.

@@

\item The third step is to read  nonduplicate blocks from each VM and output to the final backup 
storage. In this process, if a segment is not dirty, we only need to output the segment meta data 
such as reference pointers to the original segments previously saved in the backup.  
If a segment is dirty,  we use the duplicate summary information from Step 2 for internal 
duplicate blocks of this dirty segment and output the remaining non-duplicate blocks to the backup storage.
\end{itemize}

@@

The above steps can be executed by each machine
using one thread to minimize the use of CPU, network bandwidth, and disk bandwidth.
The  disk storage usage on each machine 
is fairly small for  storing part of global index and
accumulating  duplicate detection requests that contain fingerprint information.   
We impose a limit for memory allocated at each machine node to accomplish low-cost deduplication.  
Memory usage for each step of above iteration is carefully controlled as follows.
\begin{itemize}
\item For Step 1, the allocated memory for each machine is evenly divided for $p$ send buffers and $q$ receive buffers.

@@

{\bf Snapshot deleteion.} Each virtual machine will keep at most $x$ snapshots of VM images and thus
expired snapshots will be deleted unless its owern explicitly asks for its retention. 
A block or the entire segment can be deleted if its reference counter is zero.
%As we discussed in Section~\ref{sect:data}, the reference counter is kept in the global index, partitioned among machines. 
We perioidcally read all receipts of active 
snapsots and compute the counter of all blocks while still following the fingerprint-based partitioning.
That is based on the idea of mark-and-sweep~\cite{MarkSweep} while 
we deal such snapshot deletion in three partition-based steps similar to the VM snapshot backup.
Step 1 is to read  the meta data of snapshots such as segment and block
recipes and  accumate  reference count requests in different machines  in the fingerprint based distribution.
The second step  is to count references within each partition and detect those records with zero 
reference. The third step is that every VM gathers the confirmed deletion requests for its cotent and sends them
to the backup storage.
The archival data repository may log these deletion requests,  and will periodically perform a compaction operation when 
its deletion log is too big. 
Due to the paper constraint, we will not discuss the deletion operations in details.

>>>>

The data flow of our partition-based  dupliate detectionis depicted in
Figure~\ref{fig:flow}. 
%let $v$ be the total number of virtual machines hosted in  a cluster,  
%we divide the backup into $k$  iterations. 
\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{steps.pdf}
\caption{ Processing steps of dupliate detection.}
\label{fig:flow}
\end{figure}

\begin{itemize}
\item
The first step is that each machine independently reads  
%$v/k$  
virtual machine images that need a backup
and forms duplicate  detection requests. For those segments that have been modified since last backup indicated by 
dirty bits,  the system divides  each segment into a sequence of chunk blocks,  computes the meta 
information such as chunk fingerprints,  sends the metadata of a chunk block, and accumulate  into 
a partition bucket on the temporary disk storage. 
The partition mapping is using a hash function applied to the content fingerprint. 
Assuming all machines have a  homogenous resource configuration, each machine is evently  assigend with
$q$ partitons of content hash and accumulates corresponding requests on the disk. 
There are two options to assign buffers in each machine to buffer requests for each partition.
1) Each machine has  $p\times q$ send buffers corresponding to $p\times q$ partitions in the cluster
since any content block in a VM image of this machine can be sent to any of these partitions.
2) Each machine has $p$ receive buffers correspond to $p$ machines in the cluster and $q$ receive 
buffers to temporarly accmulate messages from other machines, and then output each buffer to the local disk
when it becomes full.  Option 2 is much more efficient than Option 1 because $p+q$ is much smaller than
$p\times q$, as a result each buffer has a bigger size to accmulate data and results in 
much less disk seek overhead.

@@

\item The third step is to read  nonduplicate blocks from each VM and output to the final backup 
storage. Additionally, the global index on each machine is updated with the meta data of new chunk blocks added.
In this process, if a segment is not dirty, we only need to output the segment meta data 
such as reference pointers to the original segments previously saved in the backup.  
If a segment is dirty,  we use the duplicate summary information from Step 2 for internal 
duplicate blocks of this dirty segment and output the remaining non-duplicate blocks to the backup storage.
\end{itemize}

@@

The above steps can be executed by each machine
using one thread to minimize the use of CPU, network bandwidth, and disk bandwidth.
The  disk storage usage on each machine 
is fairly small for  storing part of global index and
accumulating  duplicate detection requests that contain fingerprint information.   
We impose a limit for memory allocated at each machine node to accomplish low-cost deduplication.  
Memory usage for each step of above scheme is carefully controlled as follows.
\begin{itemize}
\item For Step 1, the allocated memory for each machine is evenly divided for $p$ send buffers and $q$ receive buffers.

@@

{\bf Snapshot deleteion.} Each virtual machine will keep at most $x$ snapshots of VM images and thus
expired snapshots will be deleted unless its owern explicitly asks for its retention. 
A block or the entire segment can be deleted if its reference counter is zero.
%As we discussed in Section~\ref{sect:data}, the reference counter is kept in the global index, partitioned among machines. 
We perioidcally read all receipts of active 
snapsots and compute the counter of all blocks while still following the fingerprint-based partitioning.
That is based on the idea of mark-and-sweep~\cite{MarkSweep} while 
we deal such snapshot deletion in three partition-based steps similar to the VM snapshot backup.
Step 1 is to read  the meta data of snapshots such as segment and block
recipes and  accumate  reference count requests in different machines  in the fingerprint based distribution.
The second step  is to count references within each partition and detect those records with zero 
reference. The third step is that every VM gathers the confirmed deletion requests for its cotent and sends them
to the backup storage.
The archival data repository may log these deletion requests,  and will periodically perform a compaction operation when 
its deletion log is too big. 
%Due to the paper constraint, we will not discuss the deletion operations in details.

### 2013-03-08 hotstorage13/eval.tex
evaluation changes

We have implemented a prototype of the parititon-based deduplication scheme on a Linux cluster
based on Alibaba's Zene cloud platform.  Objectives of our experimental evaluation are:
1) Analyze the processing time and througput of backup for a large cluster of virtual machines.
2) Assess the effectiveness  of partition-based duplication detection 
backup.  3) Examine the impacts of buffering during meta data exchange.

@@

At Alibaba's Aliyun our target is to backup cluster up to 1000 nodes with 25 VMs on each machine.
%We are running our deduplication/backup  service on 100 nodes.
%Memory usage is about 150MB space per node during backup and
%the CPU usage is very small during the experiments. 
Based on the Alibaba data studied,  each VM has about  40GB of storage  data usage on average
including OS and user data disk.  The backup of VM snapshots is completed within a few  hours every night,
and that translates to a backup throughput of 139GB per second, or 500TB per hour.
For each VM, the system keeps 10 automatically-backed snapshots in the storage while
a user may instruct extra snapshots to be saved.

@@

%Since it's impossible to perform large scale analysis without affecting the VM performance,
%We have sampled a data set from 1323 real user VMs from a cluster with 100 nodes 
%to measure the effectiveness of our scheme.
We have used a storage dataset of 1323 VM users collected from 100 Aliyun's cloud nodes to evaluate the effectiveness.
In this dataset, there are 10 snapshots per each VM user and the total amount of space investigated is 23.5 terabytes.
%This dataset contas compose of 35 VMs from 7 popular OSes: 
%Debian, Ubuntu, Redhat, CentOS, Win2003 32bit, win2003 64 bit and win2008 64 bit. For each OS, 
%5 VMs are chosen, and every VM come with 10 full snapshots of it OS and data disk. 
%The overall data size for this 700 full snapshots is 17.6 TB.

%This dataset  contains the snapshots of 1323 VMs.
%Since inner-VM deduplication is not involved in the first snapshot, this data set helps us to 
%study the CDS deduplication against user-related data. The overall size of dataset2 is 23.5 TB.
All data are divided into 2 MB fix-sized segments and each segment is divided into 
variable-sized content blocks ~\cite{similar94,rabin81} with an average size of 4KB.
The signature for variable-sized blocks is computed using their SHA-1 hash. 
%Popularity of data blocks are collected through global counting 
%and the top 1\% will fall into CDS, as discussed in Section~\ref{sect:crossVM}.

We have also used a combination of level 1 duplicate detection using segement dirty bits and
a bloom filter and prefetch cache  after level 1 basd on~\cite{Zhu}. 
The bloomer filter settings  resutls a 1:10 ratio of index reduction for in-memory search before visiting
the disk for the full index with 2\% false positives. The prefetching cache hit ratio is set to be 98\% based
a number in ~\cite{Zhu}. Our actual cache hit ratio tested is close 90\%.
\comments{

@@

\begin{verbatim}
q		50	100	250	275	300	450	500	750	1500	2000
Memory meta data 0.161	0.0805	0.0322	0.029272727	0.026833333	0.017888889	0.0161	0.010733333	0.005366667	0.004025
Global index 0.115	0.0575	0.023	0.020909091	0.019166667	0.012777778	0.0115	0.007666667	0.003833333	0.002875
Seek (Step 1) 0.0212962960.0425 0.106 0.117 0.127777778	0.191666667	0.212962963	0.319444444	0.638888889	0.851851852
Seek Option 6.388888889	12.77777778	31.94444444	35.13888889	38.33333333	57.5	63.88888889	95.83333333	191.6666667	255.5555556
Total time 3.059031718	3.078670034	3.131061779	3.137163671	3.141047877	3.313871832	3.296994302	3.377385758	3.689876824	3.901831421
throughput  per node 0.090805785	0.090226551	0.088716799	0.088544242	0.088434748	0.083822728	0.084251822	0.082246387	0.075281044	0.07119164

@@

Table~\ref{tab:overall} shows the performance of partition-based deduplication when 
the number of partitons per machine varis from 50 to 2000.
This is the overall time for a 100-node cluster and the throughput per machine reported  is the total 
amount of virtual machine images divided by the processing time. The processng time includes the first scanning time
of dirty VM segments to generate block figureprints and the second scann time for real backup.

@@

@@

>>>>

We have implemented and evaluated a prototype of the parititon-based deduplication scheme on a Linux cluster
of AMD Bulldozer FX8120 and Intel Nehalem E5530 machines.  Objectives of our experimental evaluation are:
1) Analyze the processing time and througput of backup for a large cluster of virtual machines.
2) Assess the effectiveness  of partition-based duplication detection 
backup.  3) Examine the impacts of buffering during meta data exchange.

@@

Our implementation is based on Alibaba's Xen cloud platform~\cite{Aliyun,WeiZhangIEEE}.  
At Alibaba's Aliyun, the  target is to backup cluster of a few thousands nodes with 25 VMs on each machine.
%We are running our deduplication/backup  service on 100 nodes.
%Memory usage is about 150MB space per node during backup and
%the CPU usage is very small during the experiments. 
Based on the Alibaba data studied,  each VM has about  40GB of storage  data usage on average
including OS and user data disk.  The backup of VM snapshots is completed within a few  hours every night,
%and that translates to an aggregated backup throughput of 139GB per second, or 500TB per hour.
%  For 2 hours, every machine can do 139MB/second. For 3 hours, then 92MB/second
For each VM, the system keeps 10 automatically-backed snapshots in the storage while
a user may instruct extra snapshots to be saved.

@@

%Since it's impossible to perform large scale analysis without affecting the VM performance,
%We have sampled a data set from 1323 real user VMs from a cluster with 100 nodes 
%to measure the effectiveness of our scheme.
We have performed a trace-driven study using a dataset containing 10 snapsots of 35 VMs.
That is part of a 1323 VM dataset  collected from 100 Aliyun's cloud nodes 
in our earier work with  Alibaba Aliyun~\cite{WeiZhang}.
%In this dataset, there are 10 snapshots per each VM user and the total amount of space 
%investigated is 17.5 terabytes.
%This dataset contas compose of 35 VMs from 7 popular OSes: 
%Debian, Ubuntu, Redhat, CentOS, Win2003 32bit, win2003 64 bit and win2008 64 bit. For each OS, 
%5 VMs are chosen, and every VM come with 10 full snapshots of it OS and data disk. 
%The overall data size for this 700 full snapshots is 17.6 TB.
%This dataset  contains the snapshots of 1323 VMs.
%Since inner-VM deduplication is not involved in the first snapshot, this data set helps us to 
%study the CDS deduplication against user-related data. The overall size of dataset2 is 23.5 TB.
All data are divided into 2 MB fix-sized segments and each segment is divided into 
variable-sized content blocks ~\cite{similar94,rabin81} with an average size of 4KB.
The signature for variable-sized blocks is computed using their SHA-1 hash. 
%Popularity of data blocks are collected through global counting 
%and the top 1\% will fall into CDS, as discussed in Section~\ref{sect:crossVM}.

\comments{

@@

\begin{verbatim}
memory
#partitions/node	50	100	250	275	300	450	500	750	1500	2000
Memory meta data 0.161	0.0805	0.0322	0.029272727	0.026833333	0.017888889	0.0161	0.010733333	0.005366667	0.004025
Global index 0.115	0.0575	0.023	0.020909091	0.019166667	0.012777778	0.0115	0.007666667	0.003833333	0.002875
Seek (Step 1) 0.0212962960.0425 0.106 0.117 0.127777778	0.191666667	0.212962963	0.319444444	0.638888889	0.851851852
Seek Option 2: 6.388888889	12.77777778	31.94444444	35.13888889	38.33333333	57.5	63.88888889	95.83333333	191.6666667	255.5555556
Total time 3.059031718	3.078670034	3.131061779	3.137163671	3.141047877	3.313871832	3.296994302	3.377385758	3.689876824	3.901831421
throughput  per node 0.090805785	0.090226551	0.088716799	0.088544242	0.088434748	0.083822728	0.084251822	0.082246387	0.075281044	0.07119164

@@

Table~\ref{tab:overall} shows the performance of partition-based deduplication when 
the number of partitons per machine ($q$) varies from 50 to 2000.
Row 2 shows the memory usage needed to load a partition of global index and detection requests,
varying from 161MB to 4.025MB. As we impose a constraint of 30MB, then only $q>300$ would be a viable choice.
Row 3 shows the majority of memory usage is for global index.
Row 4 is the total time for a 100-node cluster.
Row 5 is  the throughput per machine reported  is the total 
amount of virtual machine images divided by the processing time. 
The processng time includes the first scanning time
of dirty VM segments to generate block figureprints and the second scan time for real backup.

@@

Table~\ref{tab:memory} shows the performance of partition-based deduplication when 
the memory limit imposed on each node, varying from 
The overall processing time does not have a signficant reduction as we increase memory usage  20MB to 200MB.
As Row 2 shows, the time is mainly dominated by  the IO costs of two scans of VMs.
Thus we conlude that using about 30MB is sufficient to have VM backup completed in parallel within about 3.11 hours.

@@

We have also compared our pre-detection approach with another approach by combining
dirtybit segement  detection and  the data domain method~\cite{Zhu}. 
The bloomer filter setting  results a 1:10 ratio of index reduction for in-memory search before visiting
the disk for the full index with 2\% false positives. The prefetching cache hit ratio is set to be 98.96\% based
a number in ~\cite{Zhu}.
We measure its  total time including one scan of VM dirty segemnts in about 1.38 hours.
That approach takes about 3.19 hours comparable with our pre-detection scheme while its memory cost
includes 1GB of bloom filters and additional space for cache. That is an significant amount of space,
compeuting with other cloud service while memory cost of our setting is  insignifcant.

%Our actual cache hit ratio tested is close 90\%.

### 2013-03-08 hotstorage13/intro.tex
evaluation changes

Among the three factors - time, cost and de-deduplication efficiency, one of them has to be compromised for the other two. For instance, if we were build a deduplication system that has a high rate of duplication detection and has a very fast response time, 
it would need a lot of memory to hold fingerprint index and do fast comparison.  This leads to a compromise on cost. 
Our objective in this paper is to lower the cost incurred while sustaining the highest de-duplication ratio. 
%at the expense of processing latency.  
At the same time, we ensure the aggregate read/write throughput of the system 
is sufficiently scalable  for dealing with backup of a large number of virtual machine images. 
% This business model partially  resembles that of Amazon Glacier.

%Our solution  only uses a very small amount of memory space during backup while completing  snapshot backup and 
%deletion in a reasonable amount of time. 
The traditional approach to  deduplication for backup is an inline approach which follows
a sequence of VM block reading, duplicate detection,  and nonduplicate  block write to the 
backup storage.  
Our key idea  is to  perform parallel duplicate detection first for VM content blocks 
among all machines before performing actual data backup. Each machine
accumulates detection requests and  then performs detection   parititon by partition 
with minimal resource usage.
Fignureprint based partitioning allows highly parallel duplicate detection  and also simplifies 
reference counting management  during snapshot detection.
% as the entire process can also be divided into a
%parition-wise  deletion. 
With careful parallelism management and message buffering during cross-machine data exchange,
this partition-based  deduplication scheme can provide  a sufficient 
throughput for large-scale cloud data  backup.

>>>>

Among the three factors - time, cost and de-deduplication efficiency, one of them has to be compromised for the other two. For instance, if we were build a deduplication system that has a high rate of duplication detection and has a very fast response time, 
it would need a lot of memory to hold fingerprint index and do fast comparison.  This leads to a compromise on cost. 
The objective of our project is to lower the cost incurred while sustaining the highest de-duplication ratio. 
%at the expense of processing latency.  
At the same time, we ensure the aggregate read/write throughput of the system 
is sufficiently scalable  for dealing with backup of a large number of virtual machine images. 
% This business model partially  resembles that of Amazon Glacier.

%Our solution  only uses a very small amount of memory space during backup while completing  snapshot backup and 
%deletion in a reasonable amount of time. 
The traditional approach to  deduplication for backup is an inline approach which follows
a sequence of VM block reading, duplicate detection,  and nonduplicate  block write to the 
backup storage.  
Our key idea  is to  first perform parallel duplicate pre-detection for VM content blocks 
among all machines before performing actual data backup. Each machine
accumulates detection requests and  then performs pre-detection   parititon by partition 
with minimal resource usage.
Fignureprint based partitioning allows highly parallel duplicate pre-detection  and also simplifies 
reference counting management.  
% as the entire process can also be divided into a
%parition-wise  deletion. 
With careful parallelism management and message buffering during cross-machine data exchange,
this pre-detecton scheme can provide  a sufficient throughput for large-scale cloud data  backup.
